rawPaper

### 2020-10-27 analysis/paper/rawPaper.Rmd
until seqs aligned --next duplicates and TAC

# reading the sequences recovered during this paper

SycidiumCuba <- NA
                                                       # DOWNLOADING SEQS FROM GENBANK

# this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/) 
install.packages('rentrez')
library(rentrez)

# checking/setting working directory as raw_data for storing the final fasta file, which are the sequences before aligning them in the next chunk 
#getwd()
# setwd is temporal, after fishining this chunk, the wd will be again /analysis/paper
#setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")
#getwd()

@@

Awaous_ocellarisCOI <- "Awaous_ocellaris[Organism] AND COI[Gene]"
Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 2)
Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")

@@

# installing necessary package for sequence manipulation
install.packages('seqinr')
library(seqinr)

@@

Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editting species name
Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        # eliminating characters after "." using gsub
Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)       # eliminating characters before "|" using gsub

# installing necessary package
install.packages("ape")
library(ape)

@@

                                            # ALIGNNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR PHYLOGENETICS

@@

fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting those with specific name in them       # 562

@@

### Aligning the sequences

BiocManager::install("msa")
library(msa)

alignmentFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_lengths.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

length(alignmentFasta)

myClustalOmegaAlignment <- msa("alignmentFasta", type = "DNA", "ClustalOmega")

print(myClustalOmegaAlignment, showConsensus=FALSE, halfNrow=3)
print(myClustalOmegaAlignment, showNames=FALSE, show="complete")

### Removing most duplicated sequences

>>>>

                                                      # INSTALLING PACKAGES AND LIBRARIES - if needed
packages = c("rentrez", "ape", "seqinr", "BiocManager", "msa")
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

#if (!requireNamespace("BiocManager", quietly=TRUE))
#install.packages("BiocManager")
#BiocManager::install("msa")

                                                       # DOWNLOADING SEQS FROM GENBANK

# rentrez: this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/) 
library(rentrez)     # delete this line after finishing

@@

Awaous_ocellarisCOI <- "Awaous_ocellaris[Organism] AND COI[Gene]"
Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 4)
Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")

@@

# seqinr: package for sequence manipulation

library(seqinr)     # delete this line after finishing

@@

Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editing species name
Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        # eliminating characters after "." using gsub
Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)       # eliminating characters before "|" using gsub

library(ape)   # delete this line after finishing

@@

                                     # ALIGNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR THE PHYLOGENETIC ANALYSES

@@

fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562

@@

### Aligning the sequences
# using libraries "msa" and "Biostrings"

library(msa)           # delete this line after finishing
library(Biostrings)    # delete this line after finishing

mySequences <- readDNAStringSet("../data/raw_data/Sicydium_COI_lengths.fasta")

myClustalOmegaAlignment <- msa(mySequences, "ClustalOmega")   # using algorithm Clustal Omega (Sievers et al. 2011) for aligning
#print(myClustalOmegaAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")

# removing samples with to many gaps at the beginning and the end of the sequences 
colmask(myClustalOmegaAlignment) <- IRanges(start=c(1,666),end=c(13,679))                
myClustalOmegaAlignment2 <- as(myClustalOmegaAlignment, "DNAStringSet")
writeXStringSet(myClustalOmegaAlignment2, file="../data/raw_data/Sicydium_COI_alignment.fasta")         # fasta file of 652 bp

### Removing most duplicate sequences
